---
title: "Predicted seagrass carbon map at Cadiz Bay for 2024 and 2025"
author: "Lee, Chengfa Benjamin"
date: "`r Sys.Date()`"
output: bookdown::html_document2
# output: powerpoint_presentation
bibliography: references.bib 
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(magrittr)
library(sf)
library(dplyr)
library(tidyr)
library(leaflet)
library(viridis)
library(terra)
library(ggplot2)
```

This page documents the results of the _Zostera noltii_ seagrass carbon mapping work at Cadiz Bay. This includes work done in:

* [Coring Carbon Seagrass](https://sigoiry.github.io/Coring_Carbon_Seagrasses/) by Simon Oiry
* [Exploring the relationship between seagrass reflectance and biomass](https://leechengfa.github.io/Seagrass_reflectance_biomass_2025.github.io/)
* [Measuring the carbon content in _Zostera noltii_ seagrass](https://leechengfa.github.io/seagrass_carbon_2025/)


All analyses were performed in R version 4.5.2 [@Rversion].
All work belong to the [REWRITE project](https://rewriteproject.eu/).

# Study Location

This page focuses on Cadiz Bay, a [demonstrator site for the REWRITE PROJECT](https://rewriteproject.eu/demonstrator-network/cadiz-bay). There is an extensive seagrass meadow along the bay.  
<!-- Its seagrass meadow was estimated to be 3.52 km^2^ during the peak season of 2024 [@davies2024]. -->

The sites were sampled according to transect lines.

```{r samplingDesign}
#| echo: false
#| error: false
#| message: false
#| warning: false

stations <- read.csv('data/shp/Cadiz_SamplingSites.txt', sep = '\t') %>%
  st_as_sf(coords = c('X', 'Y')) %>%
  rename(name = X.1) %>%
  mutate(Cover = ifelse(X.cover == '100%', 'seagrass', ifelse(X.cover == '0%', 'non-seagrass', NA)))

# group_bounds <- read_sf('data/shp/Polygons.shp')
#   group_bounds$colour <- c('darkred','darkgreen')

map_centre <- stations %>%
  st_union() %>%
  st_centroid() %>%
  st_coordinates()

pal.shp <- colorFactor(c('darkgreen','brown4'), domain = c('seagrass','non-seagrass'))

m.shp <- leaflet(stations) %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(map_centre[1], map_centre[2], 12) %>%
  addCircleMarkers(radius = 2, color = ~pal.shp(Cover), fillOpacity = 1, label = ~name) %>%
  addLegend(pal = pal.shp, values = ~Cover, opacity = 1)

m.shp
```

# Seasonal change

```{r phenologySetUp}
#| echo: false
#| error: false
#| message: false
#| warning: false

phenology <- data.frame(list(
  "tile.name" = list.files('data/ice_creams_prediction') %>%
    gsub('_Pred_BFRD_V1_3.tif', '', .)
))
  phenology$date <- lapply(phenology$tile.name, FUN = function(string) {
    x <- string %>%
      strsplit("_") %>%
      unlist() %>%
      "["(3) %>%
      strsplit("T") %>%
      unlist() %>%
      "["(1)

    return(x)
  }) %>% unlist() %>% as.Date('%Y%m%d')

  phenology$ndvi.mean <- NA
  phenology$ndvi.sd <- NA

import.icecreams <- function(image.path) {
  map.ice <- rast(image.path)
    map.ice[map.ice == 0 | map.ice == -1] <- NA
    
  return(map.ice)
}
  
import.ndvi <- function(image.path) {
  map.index <- rast(image.path)
  
  return(map.index)
}  

for (n in phenology$tile.name) {
  pheno.icecreams <- list.files('data/ice_creams_prediction', full.names = TRUE) %>%
    grep(n, ., value = TRUE) %>%
    import.icecreams()
  
  pheno.ndvi <- list.files('data/ndvi', full.names = TRUE) %>%
    grep(n, ., value = TRUE) %>%
    import.ndvi() %>%
    mask(pheno.icecreams$Out_Class, inverse = TRUE, maskvalues = 4)
  pheno.ndvi[pheno.icecreams$Seagrass_Cover < 20] <- NA

  phenology$ndvi.mean[phenology$tile.name == n] <- global(pheno.ndvi, fun = 'mean', na.rm = TRUE)
  phenology$ndvi.sd[phenology$tile.name == n] <- global(pheno.ndvi, fun = 'sd', na.rm = TRUE)
  phenology$totalCover[phenology$tile.name == n] <- global(pheno.icecreams$Seagrass_Cover, fun = 'sum', na.rm = TRUE)
}
```

```{r phenologyndvi}
#| echo: false
#| error: false
#| message: false
#| warning: false
#| fig.width: 10
#| fig.cap: Mean NDVI value of pixels classified by the ICE CREAMS model as seagrass in 2024 and 2025. The black line denotes the actual mean NDVI values while the blue line and grey ribbons are the geom_smooth values. The dotted red line indicates the chosen date (2025-06-14) for showcasing the analyses.

phenology$ndvi.mean <- unlist(phenology$ndvi.mean)
phenology$ndvi.sd <- unlist(phenology$ndvi.sd)
phenology$totalCover <- unlist(phenology$totalCover)

phenology$year <- format(phenology$date,"%Y") %>% 
  as.character()
phenology$daymonth <- format(phenology$date,"%d-%m") %>% 
  as.character() %>%
  as.Date("%d-%m")
phenology$doy <- format(phenology$date, "%j") %>%
  as.integer()

year_scheme <- phenology$date %>% 
  format('%Y') %>%
  unique() %>%
  length() %>%
  viridis()

## abnormally low NDVI on 2 Nov 2022; unable to understand; further investigation needed
phenology <- phenology %>% 
  filter(!date == '2022-11-02')

# ggplot(phenology, aes(x = daymonth, y = ndvi.mean, colour = NA)) +
#   # geom_ribbon(aes(ymin = ndvi.mean - ndvi.sd, ymax = ndvi.mean + ndvi.sd), alpha = 0.5) +
#   geom_vline(xintercept = format(as.Date('20250614', '%Y%m%d'), "%d-%m") %>% as.character() %>% as.Date("%d-%m"), colour = 'red', linetype = 2, linewidth = 1) +
#   # geom_vline(xintercept = as.Date('20250715', '%Y%m%d'), colour = 'red', linetype = 2, linewidth = 1) +
#   geom_point(aes(colour = year)) +
#   geom_smooth(aes(colour = 'Smoothed trend')) + 
#   scale_x_date(date_breaks = 'month', date_labels = '%d %b', expand = c(0,0)) + 
#   xlab('Calendar date') + ylab('Mean NDVI over seagrass pixels') +
#   # manual legend creation 
#   scale_colour_manual('Legend', values = c('2016' = year_scheme[1], '2017' = year_scheme[2], '2018' = year_scheme[1], '2019' = year_scheme[4],
#                                            '2020' = year_scheme[5], '2021' = year_scheme[6], '2022' = year_scheme[7], '2023' = year_scheme[8],
#                                            '2024' = year_scheme[9], '2025' = year_scheme[10], 
#                                            'Smoothed trend' = 'blue')) +
#   theme(legend.position = 'bottom', legend.direction = "horizontal")

midmonth <- c("2021-01-15","2021-02-14","2021-03-16","2021-04-15","2021-05-16","2021-06-16",
              "2021-07-16","2021-08-16","2021-09-15","2021-10-16","2021-11-15","2021-12-16") %>%
  as.Date()

ggplot(phenology, aes(x = doy, y = ndvi.mean, colour = NA)) +
  geom_vline(xintercept = format(as.Date('20250614', '%Y%m%d'), "%j") %>% as.integer(), colour = 'red', linetype = 2, linewidth = 1) +
  geom_point(aes(colour = year)) +
  geom_smooth(aes(colour = 'Smoothed trend'), method = 'gam', formula = y ~ s(x, bs = 'cc', k = 10)) + 
  scale_x_continuous(breaks = format(midmonth, '%j') %>% as.integer(), labels = format(midmonth, '%B'), limits = c(1, 365)) +
  xlab('Calendar date') + ylab('Mean NDVI over seagrass pixels') +
  # manual legend creation 
  scale_colour_manual('Legend', values = c('2016' = year_scheme[1], '2017' = year_scheme[2], '2018' = year_scheme[1], '2019' = year_scheme[4],
                                           '2020' = year_scheme[5], '2021' = year_scheme[6], '2022' = year_scheme[7], '2023' = year_scheme[8],
                                           '2024' = year_scheme[9], '2025' = year_scheme[10], 
                                           'Smoothed trend' = 'blue')) +
  theme(legend.position = 'bottom', legend.direction = "horizontal")
```

```{r phenologycover}
#| echo: false
#| error: false
#| message: false
#| warning: false
#| fig.width: 10
#| fig.cap: Total seagrass extent of pixels classified by the ICE CREAMS model as seagrass in 2024 and 2025, by assuming that every percent in the seagrass cover is equivalent to 1 m^2. The black line denotes the predicted extent while the blue line and grey ribbons are the geom_smooth values. The dotted red line indicates the chosen date (2025-06-14) for showcasing the analyses.

ggplot(phenology, aes(x = doy, y = totalCover/1000000, colour = NA)) +
  geom_vline(xintercept = format(as.Date('20250614', '%Y%m%d'), "%j") %>% as.integer(), colour = 'red', linetype = 2, linewidth = 1) +
  geom_point(aes(colour = year)) +
  geom_smooth(aes(colour = 'Smoothed trend'), method = 'gam', formula = y ~ s(x, bs = 'cc', k = 10)) + 
  scale_x_continuous(breaks = format(midmonth, '%j') %>% as.integer(), labels = format(midmonth, '%B'), limits = c(1, 365)) +
  xlab('Calendar date') + ylab('Total seagrass extent (km2)') +
  # manual legend creation 
  scale_colour_manual('Legend', values = c('2016' = year_scheme[1], '2017' = year_scheme[2], '2018' = year_scheme[1], '2019' = year_scheme[4],
                                           '2020' = year_scheme[5], '2021' = year_scheme[6], '2022' = year_scheme[7], '2023' = year_scheme[8],
                                           '2024' = year_scheme[9], '2025' = year_scheme[10], 
                                           'Smoothed trend' = 'blue')) +
  theme(legend.position = 'bottom', legend.direction = "horizontal")
```

# ICE CREAMS Model prediction

For the purpose of map display and data analysis, the result from the Sentinel-2 image on 14 June 2025 was chosen. This date falls within the fieldwork dates and is thus best for validation and groundtruthing. 

```{r customleafletfunctionlegenddecreasing}
#| echo: false
#| error: false
#| message: false
#| warning: false

## function copied from stackoverflow
## https://stackoverflow.com/questions/40276569/reverse-order-in-r-leaflet-continuous-legend
addLegend_decreasing <- function (map, position = c("topright", "bottomright", "bottomleft","topleft"),
                                  pal, values, na.label = "NA", bins = 7, colors, 
                                  opacity = 0.5, labels = NULL, labFormat = labelFormat(), 
                                  title = NULL, className = "info legend", layerId = NULL, 
                                  group = NULL, data = getMapData(map), decreasing = FALSE) {
  
        position <- match.arg(position)
        type <- "unknown"
        na.color <- NULL
        extra <- NULL
        if (!missing(pal)) {
            if (!missing(colors)) 
                stop("You must provide either 'pal' or 'colors' (not both)")
            if (missing(title) && inherits(values, "formula")) 
                title <- deparse(values[[2]])
            values <- evalFormula(values, data)
            type <- attr(pal, "colorType", exact = TRUE)
            args <- attr(pal, "colorArgs", exact = TRUE)
            na.color <- args$na.color
            if (!is.null(na.color) && col2rgb(na.color, alpha = TRUE)[[4]] == 
                    0) {
                na.color <- NULL
            }
            if (type != "numeric" && !missing(bins)) 
                warning("'bins' is ignored because the palette type is not numeric")
            if (type == "numeric") {
                cuts <- if (length(bins) == 1) 
                    pretty(values, bins)
                else bins   
                if (length(bins) > 2) 
                    if (!all(abs(diff(bins, differences = 2)) <= 
                                     sqrt(.Machine$double.eps))) 
                        stop("The vector of breaks 'bins' must be equally spaced")
                n <- length(cuts)
                r <- range(values, na.rm = TRUE)
                cuts <- cuts[cuts >= r[1] & cuts <= r[2]]
                n <- length(cuts)
                p <- (cuts - r[1])/(r[2] - r[1])
                extra <- list(p_1 = p[1], p_n = p[n])
                p <- c("", paste0(100 * p, "%"), "")
                if (decreasing == TRUE){
                    colors <- pal(rev(c(r[1], cuts, r[2])))
                    labels <- rev(labFormat(type = "numeric", cuts))
                }else{
                    colors <- pal(c(r[1], cuts, r[2]))
                    labels <- rev(labFormat(type = "numeric", cuts))
                }
                colors <- paste(colors, p, sep = " ", collapse = ", ")
            }
            else if (type == "bin") {
                cuts <- args$bins
                n <- length(cuts)
                mids <- (cuts[-1] + cuts[-n])/2
                if (decreasing == TRUE){
                    colors <- pal(rev(mids))
                    labels <- rev(labFormat(type = "bin", cuts))
                }else{
                    colors <- pal(mids)
                    labels <- labFormat(type = "bin", cuts)
                }
            }
            else if (type == "quantile") {
                p <- args$probs
                n <- length(p)
                cuts <- quantile(values, probs = p, na.rm = TRUE)
                mids <- quantile(values, probs = (p[-1] + p[-n])/2, na.rm = TRUE)
                if (decreasing == TRUE){
                    colors <- pal(rev(mids))
                    labels <- rev(labFormat(type = "quantile", cuts, p))
                }else{
                    colors <- pal(mids)
                    labels <- labFormat(type = "quantile", cuts, p)
                }
            }
            else if (type == "factor") {
                v <- sort(unique(na.omit(values)))
                colors <- pal(v)
                labels <- labFormat(type = "factor", v)
                if (decreasing == TRUE){
                    colors <- pal(rev(v))
                    labels <- rev(labFormat(type = "factor", v))
                }else{
                    colors <- pal(v)
                    labels <- labFormat(type = "factor", v)
                }
            }
            else stop("Palette function not supported")
            if (!any(is.na(values))) 
                na.color <- NULL
        }
        else {
            if (length(colors) != length(labels)) 
                stop("'colors' and 'labels' must be of the same length")
        }
        legend <- list(colors = I(unname(colors)), labels = I(unname(labels)), 
                                     na_color = na.color, na_label = na.label, opacity = opacity, 
                                     position = position, type = type, title = title, extra = extra, 
                                     layerId = layerId, className = className, group = group)
        invokeMethod(map, data, "addLegend", legend)
        }
```

```{r icecreams}
#| echo: false
#| error: false
#| message: false
#| warning: false

pal.icecreams <- c('#ded4c3','#926e62','#c6c97c','#8ca271','#ddac74','#a9665e','#ca474f','#9ec8c7','#b8dc70')
labels.icecreams <- c("Bare Sand","Bare Sediment","Chlorophyta","Magnoliopsida","Microphytobenthos","Phaeophyceae","Rhodophyta","Water","Xanthophyceae")

map.class <- import.icecreams('data/ice_creams_prediction/S2B_MSIL2A_20250614T110619_N0511_R137_T29SQA_20250614T140230_Pred_BFRD_V1_3.tif')
  levels(map.class$Out_Class) <- data.frame(id=1:9, cover = labels.icecreams)

pal.class <- colorFactor(pal.icecreams, domain = 1:9, na.color = "transparent")
```

```{r icecreamsseagrass}
#| echo: false
#| error: false
#| message: false
#| warning: false

map.class.seagrass <- map.class
  map.class.seagrass$Out_Class[map.class.seagrass$Out_Class != 4] <- NA
  levels(map.class.seagrass$Out_Class) <- data.frame(id=4, cover = labels.icecreams[4])
```

```{r ndvi}
#| echo: false
#| error: false
#| message: false
#| warning: false

map.ndvi <- import.ndvi('data/ndvi/S2B_MSIL2A_20250614T110619_N0511_R137_T29SQA_20250614T140230_ndvi.tif')

pal.ndvi <- colorNumeric(c('white','darkgreen'), domain = c(0, 1), na.color = "transparent")
```

```{r icecreamsspc}
#| echo: false
#| error: false
#| message: false
#| warning: false

pal.spc <- colorNumeric(c('white','darkgreen'), domain = c(20, 100), na.color = "transparent")
```

```{r icecreamsuncertainty}
#| echo: false
#| error: false
#| message: false
#| warning: false

pal.unc <- colorNumeric(viridis_pal(option = 'D')(100), domain = c(0, 1), na.color = "transparent")
```

```{r icecreamscombinedmap}
#| echo: false
#| error: false
#| message: false
#| warning: false

m.icecreams <- leaflet() %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(map_centre[1], map_centre[2], 12) %>%
  # ice creams class map
  addRasterImage(map.class, colors = pal.class, group = 'Classification map') %>%
  addLegend(position = 'bottomleft', colors = pal.icecreams, values = 1:9, labels = labels.icecreams, opacity = 1, group = 'Legend: Class') %>%
  # ice creams seagrass only map
  addRasterImage(map.class.seagrass, colors = pal.icecreams[4], group = 'Seagrass map') %>%
  addLegend(position = 'bottomleft', colors = pal.icecreams[4], values = 4, labels = labels.icecreams[4], opacity = 1, group = 'Legend: Seagrass') %>%
  # NDVI map
  addRasterImage(map.ndvi, colors = pal.ndvi, group = 'NDVI') %>%
  addLegend_decreasing(position = 'bottomleft', pal = pal.ndvi, values = c(0.2, 1), opacity = 1, decreasing = TRUE, group = 'Legend: NDVI') %>%
  # ice creams seagrass percent cover map
  addRasterImage(map.class$Seagrass_Cover, colors = pal.spc, group = 'Seagrass percent cover') %>%
  addLegend_decreasing(position = 'bottomleft', pal = pal.spc, values = c(20, 100), opacity = 1, decreasing = TRUE, group = 'Legend: Percent cover (%)') %>%
  # ice creams uncertainty map 
  addRasterImage(map.class$Class_Probs, colors = pal.unc, group = 'Class probability') %>%
  addLegend_decreasing(position = 'bottomleft', pal = pal.unc, values = c(0, 1), opacity = 1, decreasing = TRUE, group = 'Legend: Probability') %>%
  # # outline of natural & rewilded sites
  # addPolygons(data = group_bounds, color = ~colour, opacity = 1, fillColor = 'transparent', group = 'Sample sites') %>%
  # addLegend(position = 'bottomleft', colors = c('darkred','darkgreen'), values = c('rewilded', 'natural'), labels = c('Rewilded', 'Natural'), group = 'Sample sites') %>%
  # layer control
  addLayersControl(
    baseGroups = c('Classification map', 'Seagrass map', 'NDVI', 'Seagrass percent cover', 'Class probability'),
    # overlayGroups = c('Sample sites', 'Legend: Class', 'Legend: Seagrass', 'Legend: NDVI', 'Legend: Percent cover (%)', 'Legend: Probability'),
    overlayGroups = c('Legend: Class', 'Legend: Seagrass', 'Legend: NDVI', 'Legend: Percent cover (%)', 'Legend: Probability'),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  hideGroup(c('Legend: Seagrass', 'Legend: NDVI', 'Legend: Percent cover (%)', 'Legend: Probability'))
m.icecreams
```

# Biomass prediction 

```{r predictedbiomass}
#| echo: false
#| error: false
#| message: false
#| warning: false

derive.agb <- function(ndvi) {
  agb <- ndvi * 140 - 9.5
    agb[agb < 0] <- 0
  
  return(mask(agb, map.class$Out_Class, inverse = TRUE, maskvalues = 4))
}

map.agb <- derive.agb(map.ndvi)

derive.bgb <- function(agb) {
  bgb <- agb * 0.69 + 9.1
    bgb[agb == 0] <- 0
  
  return(bgb)
}

map.bgb <- derive.bgb(map.agb)

map.tgb <- map.agb + map.bgb

## calculate AGB and BGB over unit area (g/m2) 
## conversion first to stock by multiplying by pixel area of Sentinel-2
## then conversion to kg by dividing by 100
s2map_to_stock <- function(map) {
  total <- map %>%
    global(fun = 'sum', na.rm = TRUE) %>%
    "*"(100) %>%
    unlist()
  
  return(total)
}

## conversion to Mg (divide by 1e6)
biomass.agb <- s2map_to_stock(map.agb) / 1000000
biomass.bgb <- s2map_to_stock(map.bgb) / 1000000
biomass.tgb <- s2map_to_stock(map.tgb) / 1000000

## shared palette
max.biomass <- global(map.tgb, fun = 'max', na.rm = TRUE) %>%
  unlist() %>%
  round() 
pal.biomass <- colorNumeric(c('palegreen','darkolivegreen'), domain = c(0, max.biomass), na.color = 'transparent')
```

```{r predictedcarbonstock}
#| echo: false
#| error: false
#| message: false
#| warning: false

## g/m2 * g/g --> g as output
cadiz.carbon.leaf.factor <- 0.279073 
derive.carbon.leaf <- function(map.biomass.leaf) {
  map.biomass.leaf * cadiz.carbon.leaf.factor
}

map.agc <- derive.carbon.leaf(map.agb)

## g/m2 * g/g --> g as output
cadiz.carbon.root.factor <- 0.271339 
derive.carbon.root <- function(map.biomass.root) {
  map.biomass.root * cadiz.carbon.root.factor
}

map.bgc <- derive.carbon.leaf(map.bgb)

# calculate total seagrass extent covered, based on a Sentinel-2 pixel area
area.seagrass <- map.class.seagrass$Out_Class %>%
  noNA() %>%
  s2map_to_stock()

# based on Ana Sousa's paper that the mean annual seagrass soil carbon in the top 10 cm is 162 g C/m2; mean monthly soil carbon for July 2025 in Aveiro is 150 g C/m2
sousa.soil.mean <- 150.8
carbon.soil <- sousa.soil.mean * area.seagrass / 1000000

map.tgc <- map.agc + map.bgc + sousa.soil.mean



## conversion to Mg (divide by 1e6)
carbon.agc <- s2map_to_stock(map.agc) / 1000000
carbon.bgc <- s2map_to_stock(map.bgc) / 1000000
carbon.tgc <- s2map_to_stock(map.tgc) / 1000000

## shared palette
max.carbon <- global(map.tgc, fun = 'max', na.rm = TRUE) %>%
  unlist() %>%
  round()
pal.carbon <- colorNumeric(viridis_pal(option = 'G', direction = -1)(100), domain = c(0, max.carbon), na.color = 'transparent')
```

```{r mapbiomass}
#| echo: false
#| error: false
#| message: false
#| warning: false

m.biomass <- leaflet() %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(map_centre[1], map_centre[2], 12) %>%
  # NDVI map
  addRasterImage(map.ndvi, colors = pal.ndvi, group = 'NDVI') %>%
  addLegend_decreasing(position = 'bottomleft', pal = pal.ndvi, values = c(0.2, 1), opacity = 1, decreasing = TRUE, group = 'Legend: NDVI') %>%
  # AGB map
  addRasterImage(map.agb, colors = pal.biomass, group = 'Aboveground biomass') %>%
  addLegend_decreasing(position = 'bottomleft', pal = pal.biomass, values = c(0, max.biomass), opacity = 1, decreasing = TRUE, group = 'Legend: Biomass (g DW/m<sup>2</sup>)') %>%
  # BGB map
  addRasterImage(map.bgb, colors = pal.biomass, group = 'Belowground biomass') %>%
  # TGB map
  addRasterImage(map.tgb, colors = pal.biomass, group = 'Total biomass') %>%
  # AGC map
  addRasterImage(map.agc, colors = pal.carbon, group = 'Aboveground carbon') %>%
  addLegend_decreasing(position = 'bottomleft', pal = pal.carbon, values = c(0, max.carbon), opacity = 1, decreasing = TRUE, group = 'Legend: Carbon (g C/m<sup>2</sup>)') %>%
  # BGC map
  addRasterImage(map.bgc, colors = pal.carbon, group = 'Belowground carbon') %>%
  # TGC map
  addRasterImage(map.tgc, colors = pal.carbon, group = 'Total carbon') %>%
  # # outline of natural & rewilded sites
  # addPolygons(data = group_bounds, color = ~colour, opacity = 1, fillColor = 'transparent', group = 'Sample sites') %>%
  # addLegend(position = 'bottomleft', colors = c('darkred','darkgreen'), values = c('rewilded', 'natural'), labels = c('Rewilded', 'Natural'), group = 'Sample sites') %>%
  # layer control
  addLayersControl(
    baseGroups = c('NDVI', 'Aboveground biomass', 'Belowground biomass', 'Total biomass', 'Aboveground carbon', 'Belowground carbon', 'Total carbon'),
    # overlayGroups = c('Sample sites', 'Legend: NDVI', 'Legend: Biomass (g DW/m<sup>2</sup>)', 'Legend: Carbon (g C/m<sup>2</sup>)'),
    overlayGroups = c('Legend: NDVI', 'Legend: Biomass (g DW/m<sup>2</sup>)', 'Legend: Carbon (g C/m<sup>2</sup>)'),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  hideGroup(c('NDVI', 'Legend: NDVI', 'Legend: Carbon (g C/m<sup>2</sup>)')) %>%
  showGroup('Aboveground biomass')
m.biomass
```

# Carbon stock prediction

```{r biomassCarbonTable}
#| echo: false
#| error: false
#| message: false
#| warning: false

# function to calculate mean pixel with plus/minus sd
mean_sd_pixel <- function(map) {
  mean.pixel <- map %>%
    global(fun = 'mean', na.rm = TRUE) %>%
    unlist() %>%
    round(1)
  
  sd.pixel <- map %>%
    global(fun = 'sd', na.rm = TRUE) %>%
    unlist() %>%
    round(1)
  
  return(paste(mean.pixel, '\u00B1', sd.pixel))
}

table.bc <- data.frame(
    list(
      "component" = c('Aboveground', 'Belowground', 'Soil','Total'),
      "extent" = rep(round(area.seagrass/10000, 1), 4) %>% unlist(),
      "biomass_per_area" = c(mean_sd_pixel(map.agb), mean_sd_pixel(map.bgb), 'NA', mean_sd_pixel(map.tgb)),
      "biomass" = c(round(biomass.agb, 1), round(biomass.bgb, 1), 'NA', round(biomass.tgb, 1)),
      "carbon_content" = c(round(cadiz.carbon.leaf.factor*100, 1), round(cadiz.carbon.root.factor*100, 1), 'NA', 'NA'),
      "carbon_per_area" = c(mean_sd_pixel(map.agc), mean_sd_pixel(map.bgc), sousa.soil.mean, mean_sd_pixel(map.tgc)),
      "carbon_stock" = c(round(carbon.agc, 1), round(carbon.bgc, 1), round(carbon.soil, 1), round(carbon.agc + carbon.bgc + carbon.soil, 1))
    )) 

names(table.bc) <- c("Component", "Total seagrass extent<br>(ha)", "Mean biomass per area<br>(g DW/m^2^)", "Total biomass<br>(Mg DW)", "Carbon content<br>(% g C/g DW)", "Mean carbon stock per area<br>(g C/m^2^)", "Total carbon stock<br>(Mg C)")

knitr::kable(table.bc, 
             format = "pipe",
             align = 'c',
             caption = 'Seagrass biomass and carbon at Cadiz Bay. "DW" denotes the dry weight of the seagrass samples while "C" denotes the carbon in the seagrass samples. The total carbon estimate included the aboveground carbon estimate, belowground carbon estimate and the soil carbon estimate. The soil carbon estimate was extracted from the soil carbon stock per area for July 2025 in Ria de Aveiro reported in @sousa2019, which was based on the top 10 cm sediment layer.')
```

# Seasonal change in seagrass biomass

```{r phenologyCalculations}
#| echo: false
#| error: false
#| message: false
#| warning: false

  phenology$extent <- NA
  phenology$ag_b_perArea <- NA
  phenology$bg_b_perArea <- NA
  phenology$tg_b_perArea <- NA
  phenology$ag_b_total <- NA
  phenology$bg_b_total <- NA
  phenology$tg_b_total <- NA
  phenology$ag_c_perArea <- NA
  phenology$bg_c_perArea <- NA
  phenology$tg_c_perArea <- NA
  phenology$ag_c_total <- NA
  phenology$bg_c_total <- NA
  phenology$tg_c_total <- NA
  
mean_pixel <- function(map) {
  mean.pixel <- map %>%
    global(fun = 'mean', na.rm = TRUE) %>%
    unlist() %>%
    round(2)
  
  return(mean.pixel)
}

for (n in phenology$tile.name) {
  pheno.class <- list.files('data/ice_creams_prediction', full.names = TRUE) %>%
    grep(n, ., value = TRUE) %>%
    import.icecreams()
    levels(pheno.class$Out_Class) <- data.frame(id=1:9, cover = labels.icecreams)

  pheno.class.seagrass <- pheno.class$Out_Class
    pheno.class.seagrass[pheno.class.seagrass != 4] <- NA
      levels(pheno.class.seagrass) <- data.frame(id=4, cover = labels.icecreams[4])
  phenology$extent[phenology$tile.name == n] <- pheno.class.seagrass %>%
    noNA() %>%
    s2map_to_stock()

  pheno.ndvi <- list.files('data/ndvi', full.names = TRUE) %>%
    grep(n, ., value = TRUE) %>%
    import.icecreams()

  pheno.agb <- derive.agb(pheno.ndvi)
  pheno.bgb <- derive.bgb(pheno.agb)
  pheno.tgb <- pheno.agb + pheno.bgb
  pheno.agc <- derive.carbon.leaf(pheno.agb)
  pheno.bgc <- derive.carbon.root(pheno.bgb)
  pheno.tgc <- pheno.agc + pheno.bgc + sousa.soil.mean

  phenology$ag_b_perArea[phenology$tile.name == n] <- mean_pixel(pheno.agb)
  phenology$bg_b_perArea[phenology$tile.name == n] <- mean_pixel(pheno.bgb)
  phenology$tg_b_perArea[phenology$tile.name == n] <- mean_pixel(pheno.tgb)
  phenology$ag_b_total[phenology$tile.name == n] <- s2map_to_stock(pheno.agb) / 1000000
  phenology$bg_b_total[phenology$tile.name == n] <- s2map_to_stock(pheno.bgb) / 1000000
  phenology$tg_b_total[phenology$tile.name == n] <- s2map_to_stock(pheno.tgb) / 1000000

  phenology$ag_c_perArea[phenology$tile.name == n] <- mean_pixel(pheno.agc)
  phenology$bg_c_perArea[phenology$tile.name == n] <- mean_pixel(pheno.bgc)
  phenology$tg_c_perArea[phenology$tile.name == n] <- mean_pixel(pheno.tgc)
  phenology$ag_c_total[phenology$tile.name == n] <- s2map_to_stock(pheno.agc) / 1000000
  phenology$bg_c_total[phenology$tile.name == n] <- s2map_to_stock(pheno.bgc) / 1000000
  phenology$tg_c_total[phenology$tile.name == n] <- s2map_to_stock(pheno.tgc) / 1000000
}
```

```{r phenologybiomassgraph}
#| echo: false
#| error: false
#| message: false
#| warning: false
#| fig.width: 10
#| fig.cap: Predicted biomass values of the aboveground, belowground and total components of the seagrass meadows in Cadiz Bay from 2016 to 2025.

phenology.longer <- phenology %>%
  pivot_longer(contains("_"), names_to = c('component', 'material', 'scale'), names_sep = "_")

phenology.longer$material <- factor(phenology.longer$material, labels =  c('b' = 'Biomass', 'c' = 'Carbon'))
phenology.longer$scale <- factor(phenology.longer$scale, labels =  c('perArea' = "Mean~pixel~value~(g/m^{2})", 'total' = "Total~(Mg)"))
phenology.longer$component <- factor(phenology.longer$component, labels = c('ag' = 'Aboveground', 'bg' = 'Belowground', 'tg' = 'Total'))

ggplot(phenology.longer %>% filter(material == 'Biomass'), aes(x = doy, y = value, colour = component)) +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = 'cc', k = 10)) + 
  geom_point() +
  xlab('Date') + ylab('Value (See grid labels for units)') +
  scale_colour_viridis_d('Component', labels = c('ag' = 'Aboveground', 'bg' = 'Belowground', 'tg' = 'Total'), option = "D", direction = -1) +
  # scale_x_date(breaks = '1 month', date_labels = '%d %b') + 
  scale_x_continuous(breaks = format(midmonth, '%j') %>% as.integer(), labels = format(midmonth, '%B'), limits = c(1, 365)) +
  facet_grid(scale ~ ., scales = 'free', labeller = label_parsed)
```

```{r phenologycarbongraph}
#| echo: false
#| error: false
#| message: false
#| warning: false
#| fig.width: 10
#| fig.cap: Predicted carbon values of the aboveground and belowground components of the seagrass meadows in Cadiz Bay from 2016 to 2025. The total carbon estimate included the aboveground carbon estimate, belowground carbon estimate and the soil carbon estimate. The soil carbon estimate was extracted from the soil carbon stock per area for July 2025 in Ria de Aveiro reported in @sousa2019, which was based on the top 10 cm sediment layer.

ggplot(phenology.longer %>% filter(material == 'Carbon'), aes(x = doy, y = value, colour = component)) +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = 'cc', k = 10)) + 
  geom_point() +
  xlab('Date') + ylab('Value (See grid labels for units)') +
  scale_colour_viridis_d('Component', option = "D", direction = -1) +
  # scale_x_date(breaks = '1 month', date_labels = '%d %b') + 
  scale_x_continuous(breaks = format(midmonth[c(3,6,9,12)], '%j') %>% as.integer(), labels = format(midmonth[c(3,6,9,12)], '%b'), limits = c(1, 365)) + 
  facet_wrap(scale ~ component, scales = 'free', labeller = label_parsed)
```

# Reference